---
title: "Retrieving and processing agrometeorological data from API-client sources using R software"
author: "Adrian A. Correndo, Luiz H. Moro Rosso & Ignacio A. Ciampitti"
affiliation: "Department of Agronomy, Kansas State University, Manhattan, KS-66506"
date: "01/28/2021"
contact: "correndo@ksu.edu; lhmrosso@ksu.edu; ciampitti@ksu.edu."
citation: " Correndo, Adrian A.; Moro Rosso, Luiz H.; Ciampitti, Ignacio A., 2021, 'Agrometeorological data using R-software', https://doi.org/10.7910/DVN/J9EUZU, Harvard Dataverse, V5"
---

\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = F, warning = F)

```

During the tutorial: i) we provide lines of code showing how to download daily-weather data (**Section 2**), and ii) we offer the option to generate new variables and summaries for different time intervals or periods either historical or during the cropping season (**Sections 3 to 5**). <br/>

This code was generated using R version 4.0.3 (Linux-GNU, 64-bit) and R-studio v1.2.5042. Original file is R Markdown (*rmd) with code in chunks. <br/>

\newpage

## Loading packages

```{r}
library(tidyverse) # Data wrangling
library(lubridate) # Dates operations
library(daymetr)
library(chirps)
library(nasapower)
library(vegan) # Shannon Diversity Index
```

## Input example

### Creating within R

In the next chunk we create a data-table with the required formats. Please, note that we use YYYY_MM_DD format, using "_" as separator to avoid format conflicts if data is generated in Spreadsheet software such as Excel, LibreCalc, or similar. Data will be later transformed to Date-format during the code. <br/>

The user could use either the provided example, or he/she might use it as a template to fill it out with pertinent data. Each row will represent a unique site/location, and key metadata such as lat-lon coordinates, and key dates will be represented by columns. At least, user must provided "Start" and "End" dates.

```{r}
# Coordinates of each site (site names must be unique)
# Each site is a row
# Date for each site in columns
df.site <- data.frame(ID = c('1','2','3'),
                      Crop = c('Corn','Wheat','Soy'),
                      Site = c('Scandia','Belleville','Ottawa'),
                      # Both coordinates in decimal format
                      latitude = c(39.8291,39.8158,38.5398),
                      longitude = c(-97.8458,-97.6720,-95.2446))

# Specify key dates. Typically, dates relate to phenological stages
# Each date must be a column
df.time <- data.frame(ID = c('1','2','3'),
                      # Dates as YYYY_MM_DD, using "_" to separate
                      Start = c('2002_04_25','2005_10_15','2010_05_20'),
                      Flo = c('2002_07_15','2006_04_15','2010_07_05'),
                      SeFi =c('2002_08_15','2006_05_01','2010_08_15'),
                      End = c('2002_09_30','2006_06_20','2010_10_10'))

# For historical data
df.historical <- data.frame(ID = c('1','2','3'),
                      # Dates as YYYY_MM_DD, using "_" to separate
                      Start = c('2000_01_01','2000_01_01','2000_01_01'),
                      End = c('2019_12_31','2019_12_31','2019_12_31'))

# Merging sites and dates

# Seasonal
df.input <- df.site %>% left_join(df.time)

# Historical
df.historical <- df.site %>% left_join(df.historical)

```

### Creating a `.csv` template

Here we export the example tables to "csv" format to use as templates

```{r}
write.csv(df.input, 'Example_input.csv', row.names = F, na='')
write.csv(df.historical, 'Example_input_historical.csv', row.names = F, na='')
```

### Importing a `.csv` files

Here we import your table from the csv file, and show how input tables should look like right after importing:

#### Seasonal
```{r}
path = paste0(getwd(), '/') # Current directory or any path.
# Place your file in the current working directory (getwd)

# Input seasonal data
file_input = paste0(path, 'Example_input.csv') # Change to your file

# Open seasonal file
df.input <- read.table(file_input, sep=',', header = TRUE) %>% 
  mutate_at(vars(6:ncol(.)), ~as.Date(., format='%Y_%m_%d'))

# View Seasonal
kable(df.input) %>% 
  kable_styling(latex_options = c("striped"), position = "center", font_size = 10)


```
#### Historical
```{r}
path = paste0(getwd(), '/') # Current directory or any path.
# Place your file in the current working directory (getwd)

# Input historical data
file_historical = paste0(path, 'Example_input_historical.csv') # Change to your file

# Open historical file
df.historical <- read.table(file_historical, sep=',', header = TRUE) %>% 
  mutate_at(vars(6:ncol(.)), ~as.Date(., format='%Y_%m_%d'))

# View Historical
kable(df.historical) %>% 
  kable_styling(latex_options = c("striped"), position = "center", font_size = 10)

```


\newpage

# RETRIEVING & PROCESSING DATA

During the next chunks of code we will retrieve and process the weather data from the above-mentioned sources. <br/>

**Starting dates** <br/>

If the user is interested in weather of periods PRIOR to planting, he/she can define the number of Days Prior Planting (dpp) inside each "weather.source" function. By default, dpp = 0. <br/>

**Historical weather** <br/>

If the user is interested in retrieving weather from multiple years at each location, there are two main options: i) define the years as "rows" of the input data table with Start and End dates as Jan 1$^{\text{st}}$ and Dec 31$^{\text{st}}$, respectively; or ii) define the Start date of the initial year, and End date of the final year of the series. <br/>

The example here includes a separated input for historical weather (df.historical). <br/>

**Extra variables** <br/>

Neither of the databases provide data on reference evapotranspiration ($\text{ET}_0$). However, using DAYMET and NASA-POWER, it is possible to estimate $\text{ET}_0$ using the Hargreaves and Samani approach, which only requires temperature information (Hargreaves and Samani, 1985; Raziei and Pereira, 2013). However, the $\text{ET}_{0-HS}$ equation is reported togive unreliable estimates for daily $ET0$ and therefore it should be used for 10-day periods at the shortest (Cobaner et al., 2017). Check Tutorial file. <br/>

```{r}
# Constants for ET0 (Cobaner et al., 2017)
# Solar constant
Gsc = 0.0820 # (MJ m-2 min-1)
# Radiation adjustment coefficient (Samani, 2004)
kRs = 0.17

```

\newpage

## DAYMET function

Here we download the daily-weather data from the DAYMET database, and we process it to obtain common variables of agronomic value.

```{r}

# Function
weather.daymet <- function(input, dpp=0){ input %>%
  mutate(Weather = pmap(list(ID = ID,
                             lat = latitude,
                             lon = longitude,
                             sta = Start - dpp,
                             end = End),
                        
        # Retrieving daymet data
        function(ID,lat,lon,sta,end){
                download_daymet(site = ID,
                                lat = lat, lon = lon,
                                # Extracting year
                                start = as.numeric(substr(sta,1,4)),
                                end = as.numeric(substr(end,1,4)),
                                internal = T, simplify = T)})) %>% 
  
  mutate(Weather = Weather %>% 
           
           map(~mutate(.,
                       Date = as.Date(as.numeric(yday)-1, # Day of the year
                                         origin = paste0(year,'-01-01')),
                       Year = year(Date),
                       Month = month(Date),
                       Day = mday(Date)))  %>% 
           
           map(~dplyr::select(., yday, Year, Month, Day, Date,
                              measurement,value)) %>% 
           map(~spread(., 'measurement','value'))  %>% 
           map(~rename_all(., ~c("DOY", # Date as Day of the year
                                 "Year", # Year
                                 "Month", # Month 
                                 "Day", # Day of the month
                                 "Date", # Date as normal format
                                 "DL", # Day length (sec)
                                 "PP", # Precipitation (mm)
                                 "Rad", # Radiation (W/m2)
                                 "SWE", # Snow water (kg/m2)
                                 "Tmax", # Max. temp. (degC)
                                 "Tmin", # Min. temp. (degC)
                                 "VPD"))))  %>% # Vap Pres Def (Pa)
  
  mutate(Weather = pmap(list(sta=Start-dpp,
                             end = End,data=Weather), # Requested period
                        #~filter(..3, Date>=..1 & Date<= ..2))) %>% unnest() %>%
                        function(sta, end, data){
                          filter(data, Date >= sta & Date <= end) 
                        } )) %>% unnest(cols = c(Weather)) %>% 
  
  # Converting units or adding variables
  mutate(Rad = Rad*0.000001*DL, # Radiation (W/m2 to MJ/m2)
         Tmean = (Tmax+Tmin)/2, # Mean temperature (degC),
         VPD = VPD / 1000, # VPD (Pa to kPa),
         # Data for ET0
         lat_rad = latitude*0.0174533,
         dr = 1 + 0.033*cos((2*pi/365)*DOY),
         Sd = 0.409*sin((2*pi/365)*DOY - 1.39),
         ws = acos(-tan(lat_rad)*tan(Sd)),
         Ra = (24*60)/(pi) * Gsc * dr * (ws*sin(lat_rad)*sin(Sd)+
                                  cos(lat_rad)*sin(ws)),
         ET0_HS = 0.0135 * kRs * (Ra / 2.45) * (sqrt(Tmax-Tmin)) * (Tmean + 17.8),
         DL = (DL/60)/60 # Day length (hours)
          ) %>% dplyr::select(-lat_rad,-dr,-Sd,-ws,-Ra)
}
```

### Run "weather.daymet"
```{r}
# Specify input = dataframe containing sites-data 
# Specify Days prior planting. Default is dpp = 0. Here we use dpp = 30.

weather.daymet(input = df.input, dpp = 30) -> df.weather.daymet

# Exporting data as a .csv file
write.csv(df.weather.daymet, row.names = F, na='',
          file = paste0(path, 'Output_daymet.csv'))

#View(df.weather.daymet)
```


\newpage

## NASA-POWER function

Here we download the daily-weather data from the NASA-POWER database, and we process it to obtain common variables of agronomic value. Check the Tutorial fiel for the specifics of formulae. <br/>

```{r}

weather.nasapower <- function(input, dpp=0){input %>%
  
  # Retrieving the data from nasapower
  mutate(Weather = pmap(list(ID = ID,
                             lat = latitude,
                             lon = longitude,
                             sta = Start - dpp,
                             end = End),
          function(ID,lat,lon,sta,end){
          get_power(community = "AG",
                    dates = c(sta,end),
                    lonlat = c(lon,lat),
                    temporal_average = "DAILY",
                    # Variables (see package documents)
                    pars = c("T2M_MIN", # Min. temp. (degC)
                             "T2M_MAX", # Max temp. (degC)
                             "RH2M", # Relative Humidity 2M
                             "PRECTOT", # Precipitation (mm)
                             "ALLSKY_SFC_SW_DWN"))} ) )  %>% # Radiation (MJ/m2)

  mutate(Weather = Weather %>% 
           map(~as.data.frame(.)) %>%
           # Dates operations
           map(~mutate(., yday = lubridate::yday(YYYYMMDD),
                       Year = year(YYYYMMDD),
                       Month = month(YYYYMMDD),
                       Day = mday(YYYYMMDD)))  %>%
           map(~dplyr::select(., yday,Year, Month, Day,YYYYMMDD,
                              T2M_MIN,T2M_MAX,RH2M,
                              PRECTOT,ALLSKY_SFC_SW_DWN)) %>% 
           
           map(~rename_all(., ~c("DOY", # Day of the Year
                                 "Year", # Year
                                 "Month", # Month 
                                 "Day", # Day of the month
                                 "Date", # Date
                                 "Tmin", # Min. temp. (degC)
                                 "Tmax", # Max. temp. (degC)
                                 "RH", # Relative Humidity (%)
                                 "PP", # Precipitation (mm)
                                 "Rad")))) %>% # Radiation (MJ/m2)
  
  unnest(cols = c(Weather)) %>% ungroup() %>% 
  # Converting units or adding variables
  mutate(Tmean = (Tmax+Tmin)/2, # Mean temp. (degC)
         # Nasapower does not provide VPD values
         # However, it is possible to estimate it with Temp and RH.
         es = 0.6108 * exp((17.27*Tmean) / (Tmean+237.3)),
         ea = es * (RH / 100),
         # vapour Pressure deficit (kPa)
         VPD = es - ea,
         # Data for ET0
         lat_rad = latitude*0.0174533,
         dr = 1 + 0.033*cos((2*pi/365)*DOY),
         Sd = 0.409*sin((2*pi/365)*DOY - 1.39),
         ws = acos(-tan(lat_rad)*tan(Sd)),
         Ra = (24*60)/(pi) * Gsc * dr * (ws*sin(lat_rad)*sin(Sd)+
                                  cos(lat_rad)*sin(ws)),
         ET0_HS = 0.0135 * kRs * (Ra / 2.45) * (sqrt(Tmax-Tmin)) * (Tmean + 17.8)
         ) %>% dplyr::select(-es,-ea,-lat_rad,-dr,-Sd,-ws,-Ra)
}
```

### Run "weather.nasapower"
```{r}
# Specify input = dataframe containing sites-data 
# Specify Days prior planting. Default is dpp = 0. Here we use dpp = 30.

weather.nasapower(input = df.input, dpp = 30) -> df.weather.nasapower

# Exporting data as a .csv file
write.csv(df.weather.nasapower, row.names = F, na='',
          file = paste0(path, 'Output_nasapower.csv'))

View(df.weather.nasapower)
```

\newpage

## CHIRPS function

Here we download the daily-weather data from the CHIRPS database, and we process it to obtain common variables of agronomic value.
```{r}

weather.chirps <- function(input, dpp=0){ input %>%
  
  # Retrieving the data from CHIRPS
  mutate(Weather = pmap(list(ID = ID,
                             lat = latitude,
                             lon = longitude,
                             sta = Start - dpp,
                             end = End),
                        function(ID,lat,lon,sta,end){
        get_chirps(data.frame(lon = c(lon), lat = c(lat)),
                   c(as.character(sta),as.character(end)))}) ,
  # Get prec. indices
        Indices = Weather %>% map(~precip_indices(., timeseries = TRUE, 
                                                   intervals = 30))) %>% 
  
  # Organizing dataframe
  mutate(Weather = Weather %>% 
           map(~as.data.frame(.)) %>% 
           map(~dplyr::select(., date,chirps)) %>%
           # Dates operations
           map(~mutate(., yday = lubridate::yday(date),
                       Year = year(date),
                       Month = month(date),
                       Day = mday(date))) %>% 
           map(~dplyr::select(., yday, Year, Month, Day, date,chirps)) %>%
           map(~rename_all(., ~c("DOY", "Year", "Month", "Day","Date", "PP"))),
         
         Indices = Indices %>% 
           map(~as.data.frame(.)) %>% 
           map(~spread(., 'index','value')) %>% 
           map(~dplyr::select(., -id,-lon,-lat)) %>% 
           map(~rename(., Date = date))) %>% 
  
  mutate(Full = map2(.x=Weather, .y=Indices, ~left_join(.x,.y))) %>% 
  dplyr::select(-Weather, -Indices) %>% unnest(cols = c(Full))
  
}
```

### Run "weather.chirps"
```{r}
# Specify input = dataframe containing sites-data 
# Specify Days prior planting. Default is dpp = 0. Here we use dpp = 30.

weather.chirps(input = df.input, dpp = 30) -> df.weather.chirps

# Exporting data as a .csv file
write.csv(df.weather.chirps, row.names = F, na='',
          file = paste0(path, 'Output_chirps.csv'))

View(df.weather.chirps)

```


\newpage

# TIME INTERVALS

In this section we create time intervals during the cropping season using pre-specified dates as columns at the initial data table with site information. <br/>

The user can apply: i) a unique seasonal interval (season), ii) even intervals (even), or iii) customized intervales (custom). <br/>

## FULL SEASON interval
```{r}

season <- df.input %>% 
  mutate(Intervals = # Create
           map2(.x=Start,.y=End,
                ~data.frame( # New data
                  Interval = c("Season"),
                  Start.in = c(.x),
                  End.in = c(.y) ) )) %>% 
  
  dplyr::select(ID,Site,Intervals) %>% unnest(cols = c(Intervals))

kable(season) %>% 
  kable_styling(latex_options = c("striped"), position = "center", font_size = 10)

```

## EVEN intervals
``` {r}
n = 4 # Number of intervals
dpp = 30 # Days prior planting

even <- df.input %>% 
  mutate(Intervals = # Create
      map2(.x=Start,.y=End,
      ~data.frame( # New data
      Interval = c("Prev",LETTERS[1:n+1]),
      Start.in = c(.x-dpp, seq.Date(.x,.y+1,length.out=n+1)[1:n] ),
      End.in = c(.x-1,seq.Date(.x,.y+1,length.out=n+1)[2:(n+1)]))) ) %>% 
  
  dplyr::select(ID,Site,Intervals) %>% unnest(cols = c(Intervals))

kable(even) %>% 
  kable_styling(latex_options = c("striped"), position = "center", font_size = 10)

```

## CUSTOM intervals
``` {r }
# Counting # intervals
i = ncol(df.input[,6:ncol(df.input)]) # Number of intervals

# 
df.input = df.input %>% 
  # Reformat Reference dates for operations
  # Modify names and Number of dates as needed
  # Here we follow the example of df.input
  # with 4 dates named as Start (Plant), Flo, SeFi, & End
  mutate_at(vars(6:ncol(.)), 
            ~str_replace_all(as.character(.), '-','_')) %>% 
  mutate_at(vars(6:ncol(.)), ~as.Date(., format='%Y_%m_%d')) %>% data.frame() 


custom <- df.input %>% 
  mutate(Intervals = # Create
           pmap(list(x = Start - dpp,
                     y = Start,
                     z = Flo,
                     m = SeFi,
                     k = End),
                function(x,y,z,m,k){
                  data.frame( # New data
                  Interval = c(LETTERS[1:i]),
                  Name = c("Prev", "Plant-Flo","Flo-SeFi","SeFi-End"),
                  Start.in = c(x,y,z,m),
                  End.in = c(y-1,z-1,m-1,k) ) } ) ) %>% 
  
  dplyr::select(ID,Site,Intervals) %>% unnest(cols = c(Intervals))

kable(custom) %>% 
  kable_styling(latex_options = c("striped"), position = "center", font_size = 10)
```


\newpage

# SEASONAL SUMMARIES

For each of the period or interval of interest a variety of variables can be created. Here, we present a set of variables that can capture environmental variations that might be missing by analyzing standard weather data (precipitations, temperature, radiation). These variables represent an example that was used for studying influence of weather in corn yields by Correndo et al. (2021). Check Table 2 of Tutorial file. <br/>

\newpage

## Summary function - DAYMET & NASA-POWER
```{r}

# Defining the function to summarize DAYMET and/or NASA-POWER
summary.daymet.nasapower <- function(input, intervals) {

  intervals %>% 
  
  # Merging weather data
  left_join(input %>% 
              # Nesting weather data back for each site-ID
  dplyr::select_if(
    names(.) %in% c("ID", "Crop", "Site", "Date","DL", "PP",
                    "Rad", "Tmax", "Tmin", "Tmean", "VPD", "ET0_HS")) %>% 
  group_by(ID,Crop,Site) %>% nest(.key = 'Weather') %>% ungroup()) %>% 
  
  mutate(Weather = pmap(list(x = Start.in,y = End.in, data = Weather),
                        function(x, y, data){
                          filter(data, Date >= x & Date < y)} ) )  %>% 
  
  mutate(Weather = Weather %>% # User must addapt depending on the crop
           map(~mutate(.,
                # Ext. Prec. event
                EPEi = case_when(PP>25~1, TRUE~0),
                # Ext.Temp. event
                ETEi = case_when(Tmax >= 30~1, TRUE~0), 
                # Tmax factor,  crop heat units (CHU)
                Ymax = case_when(Tmax < 10~0, 
                            TRUE ~ 3.33*(Tmax-10)-0.084*(Tmax-10)),
                # Tmin factor, Crop heat units (CHU)
                Ymin = case_when(Tmin<4.44~0, 
                            TRUE ~ 1.8*(Tmin-4.44)), 
                # Daily CHU
                Yavg = (Ymax+Ymin)/2,
                # Tmin threshold Growing Degrees.
                Gmin = case_when(Tmin >= 10 ~ Tmin, 
                            TRUE ~ 10),
                # Tmax threshold Growing Degrees.
                Gmax = case_when(Tmax <= 30 ~ Tmax, 
                            TRUE ~ 30),
                # Daily Growing Degree Units.
                GDU = ((Gmin + Gmax)/2) - 10
                       ) ) ) %>% 

# Summary for each variable

  mutate(# Duration of interval (days)
         Dur = Weather %>% map(~nrow(.)),
         # Accumulated PP (mm)
         PP = Weather %>% map(~sum(.$PP)),
         # Mean Temp (C)
         Tmean = Weather %>% map(~mean(.$Tmean)),
         # Accumulated Rad (MJ/m2)
         Rad = Weather %>% map(~sum(.$Rad)),
         # Accumulated VPD (kPa)
         VPD = Weather %>% map(~sum(.$VPD)),
         # Accumulated ET0 (mm)
         ET0_HS = Weather %>% map(~sum(.$ET0_HS)),
         # Number of ETE (#)
         ETE = Weather %>% map(~sum(.$ETEi)),
         # Number of EPE (#)
         EPE = Weather %>% map(~sum(.$EPEi)),
         # Accumulated Crop Heat Units (CHU)
         CHU = Weather %>% map(~sum(.$Yavg)),
         # Shannon Diversity Index for PP
         SDI = Weather %>% map(~diversity(.$PP, index="shannon")/
                                 log(length(.$PP))),
         # Accumulated Growing Degree Days (GDD) 
         GDD =  Weather %>% map(~sum(.$GDU))) %>% 
  
# Additional indices and final units
  dplyr::select(-Weather) %>% unnest() %>% 
  mutate(# Photo-thermal quotient (Q)
         Q_chu = Rad/CHU,
         Q_gdd = Rad/GDD,
         # Abundant and Well Distributed Water
         AWDR = PP*SDI) 
}
```

\newpage

## Summary function - CHIRPS. <br/>
``` {r war}
#########

# Defining function to summarize CHIRPS data
summary.chirps <- function(input, intervals) {

intervals %>% 
  
  # Merging weather data
  left_join(input %>% 
              # Nesting weather data back for each site-ID
  dplyr::select(c(ID, Crop, Site, Date, PP)) %>% 
  group_by(ID,Crop,Site) %>% nest(.key = 'Weather') %>% ungroup()) %>% 
  
  mutate(Weather = pmap(list(x = Start.in,y = End.in, data = Weather),
                        function(x, y, data){
                          filter(data, Date >= x & Date < y)} ) )  %>% 
  
  mutate(Weather = Weather %>% # User must addapt depending on the crop
           map(~mutate(., EPEi = case_when(PP>25~1, TRUE~0) # Ext. Prec. event
                       ) ) )   %>%

# Summary for each variable
  
  mutate(# Duration of interval (days)
         Dur = Weather %>% map(~nrow(.)),
         # Accumulated PP (mm)
         PP = Weather %>% map(~sum(.$PP)),
         # Number of EPE (#)
         EPE = Weather %>% map(~sum(.$EPEi)) ,
         # Shannon Diversity Index for precipitation data
         SDI = Weather %>% map(~diversity(.$PP, index="shannon")/
                                 log(length(.$PP)))) %>% 
  
# Additional indices and final units
  dplyr::select(-Weather) %>% unnest() %>% 
  mutate(AWDR = PP*SDI) # Abundant and Well Distributed Water

}

```

\newpage

## DAYMET summary
```{r}

# Run the summary
# input = dataframe containing the data (from daymet or nasapower).
# intervals = type of intervals (season, custom or even)

df.summary.daymet <-
  summary.daymet.nasapower(input = df.weather.daymet,
                           intervals = custom)

kbl(df.summary.daymet) %>%
  kable_styling(font_size = 7, position = "center", latex_options = c("scale_down"))


# Exporting data as a .csv file
# Daymet
write.csv(df.summary.daymet, row.names = F, na='',
          file = paste0(path, 'Summary_daymet.csv'))

```

\newpage

## NASA-POWER summary
```{r}

# Run the summary
# data = dataframe containing the data (from daymet or nasapower).
# intervals = type of intervals (season, custom or even)
df.summary.nasapower <-
  summary.daymet.nasapower(input = df.weather.nasapower,
                           intervals = custom)

kbl(df.summary.nasapower) %>%
  kable_styling(font_size = 7, position = "center", latex_options = c("scale_down"))

# Exporting data as a .csv file
write.csv(df.summary.nasapower, row.names = F, na='',
          file = paste0(path, 'Summary_nasapower.csv')) 

```

\newpage

## CHIRPS summary
```{r}

# Run the summary
# data = dataframe containing the data.
# intervals = type of intervals (season, custom or even)

df.summary.chirps <-
  summary.chirps(input = df.weather.chirps,
                 intervals = custom)

kbl(df.summary.chirps) %>%
  kable_styling(font_size = 7, position = "center", latex_options = c("scale_down"))

# Exporting data as a .csv file
write.csv(df.summary.chirps, row.names = F, na='',
          file = paste0(path, 'Summary_chirps.csv')) 
```
\newpage

# HISTORICAL WEATHER

## Historical "weather.daymet"

For retrieving historical weather, user must specify the input containing the historical target dates by site.
```{r}
# Specify input = dataframe containing historical dates from sites 

weather.daymet(input = df.historical) -> hist.weather.daymet

# Exporting data as a .csv file
write.csv(hist.weather.daymet, row.names = F, na='',
          file = paste0(path, 'Hist_output_daymet.csv'))

#View(hist.weather.daymet)
```

## Historical "weather.nasapower"

```{r}
# Specify input = dataframe containing historical dates from sites 

weather.nasapower(input = df.historical) -> hist.weather.nasapower

# Exporting data as a .csv file
write.csv(hist.weather.nasapower, row.names = F, na='',
          file = paste0(path, 'Hist_output_nasapower.csv'))

#View(hist.weather.nasapower)
```

## Historical "weather.chirps"

```{r}
# Specify input = dataframe containing historical dates from sites 

weather.chirps(input = df.historical) -> hist.weather.chirps

# Exporting data as a .csv file
write.csv(hist.weather.chirps, row.names = F, na='',
          file = paste0(path, 'Hist_output_chirps.csv'))

#View(hist.weather.chirps)
```

## Intervals functions

```{r}

historical.years <- function(hist.data) {
  
# By year
hist.data %>% group_by(ID,Year) %>% 
  dplyr::select(ID, Crop, Site, latitude, longitude, Start, End, Date, Year, Month) %>% 
  summarise(Start.in = min(Date),
            End.in = max(Date))

}

historical.years.months <- function(hist.data) {
  
# By year*month
hist.data %>% group_by(ID, Year, Month) %>% 
  dplyr::select(ID, Crop, Site, latitude, longitude, Start, End, Date, Year, Month) %>% 
  summarise(Start.in = min(Date),
            End.in = max(Date))

}

```

## DAYMET Historical summary

Summary can be obtained by years or by years.months. User must specify this option at the "intervals" argument of the summary function.  <br/>

### Intervals
```{r}

# Specify hist.data = dataframe containing the historical weather data to summarize

years = historical.years(hist.data = hist.weather.daymet)

# Specify hist.data = dataframe containing the historical weather data to summarize
years.months = historical.years.months(hist.data = hist.weather.daymet)

```

### Summary

```{r}
# input = dataframe containing the historical weather data.
# intervals = type of historical intervals (years, years.months)

historical.summary.daymet <-
  summary.daymet.nasapower(input = hist.weather.daymet,
                           intervals = years)

kbl(historical.summary.daymet) %>%
  kable_styling(font_size = 7, position = "center", latex_options = c("scale_down"))

# Exporting data as a .csv file
# Daymet
write.csv(historical.summary.daymet, row.names = F, na='',
          file = paste0(path, 'Historical_summary_daymet.csv'))

```

## NASA-POWER Historical summary

Summary can be obtained by years or by years.months. User must specify this option at the "intervals" argument of the summary function. <br/>

### Intervals

```{r}

# Specify hist.data = dataframe containing the historical weather data to summarize

years = historical.years(hist.data = hist.weather.nasapower)

# Specify hist.data = dataframe containing the historical weather data to summarize
years.months = historical.years.months(hist.data = hist.weather.nasapower)

# Run the summary
# input = dataframe containing the historical weather data.
# intervals = type of historical intervals (years, years.months)
```

### Summary
```{r}
# input = dataframe containing the historical weather data.
# intervals = type of historical intervals (years, years.months)

historical.summary.nasapower <-
  summary.daymet.nasapower(input = hist.weather.nasapower,
                           intervals = years)

kbl(historical.summary.nasapower) %>%
  kable_styling(font_size = 7, position = "center", latex_options = c("scale_down"))


# Exporting data as a .csv file
# Daymet
write.csv(historical.summary.nasapower, row.names = F, na='',
          file = paste0(path, 'Historical_summary_nasapower.csv'))

```


## CHIRPS Historical summary
Summary can be obtained by years or by years.months. User must specify this option at the "intervals" argument of the summary function. <br/>

### Intervals
```{r}

# Specify hist.data = dataframe containing the historical weather data to summarize

years = historical.years(hist.data = hist.weather.chirps)

# Specify hist.data = dataframe containing the historical weather data to summarize
years.months = historical.years.months(hist.data = hist.weather.chirps)

# Run the summary
# input = dataframe containing the historical weather data.
# intervals = type of historical intervals (years, years.months)
```

### Summary
```{r}
# input = dataframe containing the historical weather data.
# intervals = type of historical intervals (years, years.months)

historical.summary.chirps <-
  summary.chirps(input = hist.weather.chirps,
                           intervals = years)

kbl(historical.summary.chirps) %>%
  kable_styling(font_size = 7, position = "center", latex_options = c("scale_down"))


# Exporting data as a .csv file
# Daymet
write.csv(historical.summary.chirps, row.names = F, na='',
          file = paste0(path, 'Historical_summary_chirps.csv'))

```
\newpage

# REFERENCES

Allen, R.G., L.S. Pereira, D. Raes, M. Smith. 1998. Crop evapotranspiration - Guidelines for computing crop water requirements. *FAO Irrigation and drainage*, 56. FAO - Food and Agriculture Organization of the United Nations. Rome, Italy. ISBN 92-5-104219-5. http://www.fao.org/3/x0490e/x0490e00.htm#Contents

Bannayan, M., Hoogenboom, G., & Crout, N.M.J., 2004. Photothermal impact on maize performance: a simulation approach. *Ecol. Modell.*, 180 (2-3), 277-290. https://doi.org/10.1016/j.ecolmodel.2004.04.022 <br/>

Bootsma, A., S. Gameda, & D.W. McKenney. 2005. Potential impacts of cli-
mate change on corn, soybeans and barley yields in Atlantic Canada. *Can J. Soil Sci.* 85:345–357. https://doi.org/10.4141/S04-025 <br/>

Butler, E.E., & Huybers, P. 2013. Adaptation of US maize to temperature variations. *Nat. Clim. Chang*. 3N, 68–72. https://doi.org/10.1038/nclimate1585 <br/>

Cobaner M., H. Citakoğlu, T. Haktanir, & O. Kisi. 2017. Modifying Hargreaves–Samani equation with meteorological variables for estimation of reference evapotranspiration in Turkey. *Hydrol. Res.* 1 April 2017, 48 (2), 480–497. https://doi.org/10.2166/nh.2016.217 <br/>

Correndo, A.A., J.L. Rotundo, N. Tremblay, S. Archontoulis, J.A. Coulter, D. Ruiz-Diaz, D. Franzen, A.J. Franzluebbers, E. Nafziger, R. Schwalbert, K. Steinke, J. Williams, C.D. Messina, & I.A. Ciampitti. 2021. Assessing the uncertainty of maize yield without nitrogen fertilization. *Field Crops Res.* 260, 2021, 107985. https://doi.org/10.1016/j.fcr.2020.107985. <br/>

Funk C., P. Peterson, M. Landsfeld, D. Pedreros, J. Verdin, S. Shukla, & J. Michaelsen. 2015. The climate hazards infrared precipitation with stations—a new environmental record for monitoring extremes. *Scientific Data* 2, 150066. https://doi.org/10.1038/sdata.2015.66. <br/>

Gilmore, E.C., & Rogers, J.S. 1958. Heat units as a method of measuring maturity in corn. *Agron. J.* 50:611–615. https://doi.org/10.2134/agronj1958.00021962005000100014x <br/>

Hargreaves,G.H., & Z.A. Samani. 1985. Reference crop evapotranspiration from temperature. *Appl. Eng. Agric.* 1(2),96–99. https://doi.org/10.13031/2013.26773 <br/>

Raziei, T., & L.S. Pereira. 2013. Estimation of ET0 with Hargreaves–Samani and FAO-PM temperature methods for a wide range of climates in Iran. *Agric. Water Manag.* 121 (2013), 1-18. https://doi.org/10.1016/j.agwat.2012.12.019 <br/>

Sparks, A. 2018. nasapower: A NASA POWER Global Meteorology, Surface Solar ENergy and Climatology Data Client for R. *J. of Open Source Softw.*, 3(30), 1035. https://doi.org/10.21105/joss.01035 <br/>

Thornton, P.E., M. Thornton, B. Mayer, Y. Wei, R. Devarakonda, R. Vose, & R.B. Cook.  2019. Daymet: daily surface weather data on a 1-km Grid for North America, Version3. ORNL DAAC, Oak Ridge, Tennessee, USA. https://daymet.ornl.gov/ <br/>

Tremblay, N., Bouroubi, Y.M., Bélec, C., Mullen, R.W., Kitchen, N.R., Thomason, W.E., Ebelhar, S., Mengel, D.B., Raun, W.R., Francis, D.D., Vories, E.D., & Ortiz‐Monasterio, I., 2012. Corn response to nitrogen is influenced by soil texture and weather. *Agron. J.*, 104,  1658-1671. https://doi.org/10.2134/agronj2012.0184 <br/>

Ye, Q., Lin, X., Adee, E., Min, D., Assefa Mulisa, Y., O'Brien, D., & Ciampitti, I.A., 2017. Evaluation of climatic variables as yield‐limiting factors for maize in Kansas. *Int. J. Climatol.* 37.S1,  464-75. https://doi.org/10.1002/joc.5015 <br/>