---
title: "30 Year Normal Weather"
author: "Kolby Grint"
date: "2/16/2021"
output: pdf_document
---

## Loading packages

```{r}
library(easypackages)
libraries('tidyverse') # Data wrangling
libraries('lubridate') # Dates operations
libraries('kableExtra') # Table formatting
libraries('daymetr','chirps','nasapower') # Weather databases
libraries('vegan') # Shannon Diversity Index

```


```{r}
# Coordinates of each site (site names must be unique)
# Each site is a row
# Date for each site in columns
df.site <- tibble(ID = c('1','2','3'), # use tibble instead data.frame (same output - tibble better)
                      Crop = c('Corn/Soy','Corn/Soy','Corn/Soy'),
                      Site = c('Arlington','Lancaster','Rock County'),
                      # Both coordinates in decimal format
                      latitude = c( 43.3091,42.8283,42.7255),
                      longitude = c(-89.3473,-90.7908,-89.0228))

# Specify key dates. Typically, dates relate to phenological stages
# Each date must be a column
df.time <- tibble(ID = c('1','2','3'),
                      # Dates as YYYY_MM_DD, using "_" to separate
                      Start = c('1988_01_01','1988_01_01','1988_01_01'),
                      #Flo = c('2002_07_15','2006_04_15','2010_07_05'),
                      #SeFi =c('2002_08_15','2006_05_01','2010_08_15'),
                      End = c('2018_12_31','2018_12_31','2018_12_31'))

# For historical data
df.historical <- tibble(ID = c('1','2','3'),
                      # Dates as YYYY_MM_DD, using "_" to separate
                      Start = c('1988_01_01','1988_01_01','1988_01_01'),
                      End = c('2018_12_31','2018_12_31','2018_12_31'))

# Merging sites and dates

# Seasonal
df.input <- df.site %>% left_join(df.time)

# Historical
df.historical <- df.site %>% left_join(df.historical)

```

### Creating a `.csv` template

Here we export the example tables to "csv" format to use as templates

```{r}
write.csv(df.input, 'Example_input.csv', row.names = F, na='')
write.csv(df.historical, 'Example_input_historical.csv', row.names = F, na='')
```

### Importing a `.csv` files

Here we import your table from the csv file, and show how input tables should look like right after importing:

#### Seasonal
```{r}
path = paste0(getwd(), '/') # Current directory or any path.
# Place your file in the current working directory (getwd)

# Input seasonal data
file_input = paste0(path, 'Example_input.csv') # Change to your file

# Open seasonal file
df.input <- read.table(file_input, sep=',', header = TRUE) %>% 
  mutate_at(vars(6:ncol(.)), ~as.Date(., format='%Y_%m_%d'))

# View Seasonal
kable(df.input) %>% 
  kable_styling(latex_options = c("striped"), position = "center", font_size = 10)


```

#### Historical
```{r}
path = paste0(getwd(), '/') # Current directory or any path.
# Place your file in the current working directory (getwd)

# Input historical data
file_historical = paste0(path, 'Example_input_historical.csv') # Change to your file

# Open historical file
df.historical <- read.table(file_historical, sep=',', header = TRUE) %>% 
  mutate_at(vars(6:ncol(.)), ~as.Date(., format='%Y_%m_%d'))

# View Historical
kable(df.historical) %>% 
  kable_styling(latex_options = c("striped"), position = "center", font_size = 10)

```

```{r}
# Constants for ET0 (Cobaner et al., 2017)
# Solar constant
Gsc = 0.0820 # (MJ m-2 min-1)
# Radiation adjustment coefficient (Samani, 2004)
kRs = 0.17
```

## DAYMET function

Here we download the daily-weather data from the DAYMET database, and we process it to obtain common variables of agronomic value.

```{r}

# Function
weather.daymet <- function(input, dpp=0){ input %>%
  mutate(Weather = pmap(list(ID = ID,
                             lat = latitude,
                             lon = longitude,
                             sta = Start - dpp,
                             end = End),
                        
        # Retrieving daymet data
        function(ID,lat,lon,sta,end){
                download_daymet(site = ID,
                                lat = lat, lon = lon,
                                # Extracting year
                                start = as.numeric(substr(sta,1,4)),
                                end = as.numeric(substr(end,1,4)),
                                internal = T, simplify = T)})) %>% 
  
  mutate(Weather = Weather %>% 
           
           map(~mutate(.,
                       Date = as.Date(as.numeric(yday)-1, # Day of the year
                                         origin = paste0(year,'-01-01')),
                       Year = year(Date),
                       Month = month(Date),
                       Day = mday(Date)))  %>% 
           
           map(~dplyr::select(., yday, Year, Month, Day, Date,
                              measurement,value)) %>% 
           map(~spread(., 'measurement','value'))  %>% 
           map(~rename_all(., ~c("DOY", # Date as Day of the year
                                 "Year", # Year
                                 "Month", # Month 
                                 "Day", # Day of the month
                                 "Date", # Date as normal format
                                 "DL", # Day length (sec)
                                 "PP", # Precipitation (mm)
                                 "Rad", # Radiation (W/m2)
                                 "SWE", # Snow water (kg/m2)
                                 "Tmax", # Max. temp. (degC)
                                 "Tmin", # Min. temp. (degC)
                                 "VPD"))))  %>% # Vap Pres Def (Pa)
  
  mutate(Weather = pmap(list(sta=Start-dpp,
                             end = End,data=Weather), # Requested period
                        #~filter(..3, Date>=..1 & Date<= ..2))) %>% unnest() %>%
                        function(sta, end, data){
                          filter(data, Date >= sta & Date <= end) 
                        } )) %>% unnest(cols = c(Weather)) %>% 
  
  # Converting units or adding variables
  mutate(Rad = Rad*0.000001*DL, # Radiation (W/m2 to MJ/m2)
         Tmean = (Tmax+Tmin)/2, # Mean temperature (degC),
         VPD = VPD / 1000, # VPD (Pa to kPa),
         # Data for ET0
         lat_rad = latitude*0.0174533,
         dr = 1 + 0.033*cos((2*pi/365)*DOY),
         Sd = 0.409*sin((2*pi/365)*DOY - 1.39),
         ws = acos(-tan(lat_rad)*tan(Sd)),
         Ra = (24*60)/(pi) * Gsc * dr * (ws*sin(lat_rad)*sin(Sd)+
                                  cos(lat_rad)*sin(ws)),
         ET0_HS = 0.0135 * kRs * (Ra / 2.45) * (sqrt(Tmax-Tmin)) * (Tmean + 17.8),
         DL = (DL/60)/60 # Day length (hours)
          ) %>% dplyr::select(-lat_rad,-dr,-Sd,-ws,-Ra)
}
```

### Run "weather.daymet"
```{r}
# Specify input = dataframe containing sites-data 
# Specify Days prior planting. Default is dpp = 0. Here we use dpp = 30.

weather.daymet(input = df.input, dpp = 0) -> df.weather.daymet

# Exporting data as a .csv file
#write.csv(df.weather.daymet, row.names = F, na='',
          #file = paste0(path, 'Output_daymet.csv'))

#View(df.weather.daymet)
```
I think the weather data starts one month early? Filtered it out.

### Kolby's Wrangling of Daymet Data
```{r Wrangling Daymet Data}
df.weather.daymet %>%
  filter(!Year == "1987")

arlweather <- df.weather.daymet %>%
  filter(Site == "Arlington") 
  

#average monthly precipitation and temperature
arlweather_monthly <- arlweather %>%
  group_by(Month) %>%
  summarise(prcp_mm0 = sum(PP)/30 , temp_C = mean(Tmean))
  #monthly precipitation is wrong. need to think more about how to do this with a fresh mind.
```




### Kolby's Solo Attempt Ignore for now and probably delete.
```{r data}
arlington<-download_daymet(site= "arlington",
                lat = 43.3091, #The N-S latitude value
                lon = -89.3473, #The E-W longitude value
                start = 1988,#Start of the range of years
                end = 2018,#end of the range of years
                path = ,#Path for where to save the data; default is NULL
                internal = TRUE,
                simplify = TRUE)

str(arlington)
```


```{r data wrangling}
arlington_prcp= arlington %>%
  select(site, tile, year, yday, measurement, value) %>%
  filter(measurement == "prcp..mm.day.") #%>%
  ##mutate(month)

arlington_tmax= arlington %>%
  select(site, tile, year, yday, measurement, value) %>%
  filter(measurement == "tmax..deg.c.") %>%
  rename( c("tmax" = "measurement", "tmaxc" = "value"))
 

arlington_tavg= arlington %>%
  select(site, tile, year, yday, measurement, value) %>%
  filter(measurement == "tmin..deg.c.") %>%
  left_join(arlington_tmax, by = c("site", "tile", "year", "yday")) %>%
  mutate(tmean = (value + tmaxc)/2)

#Probably would have worked but Couldn't 
```
Probably would have worked but Couldn't figure out how to get months assigned with only "day of year". Ultimately was trying to develop monthly normal data.

