---
title: "weather_max"
author: "Maxwel Coura Oliveira"
date: "2/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = TRUE,
  warning = TRUE)
```

```{r}
library(tidyverse) # Data wrangling
library(lubridate) # Dates operations
library(daymetr)
library(chirps)
library(nasapower)
library(vegan)
```

# Load data

You can load your data with read_csv if you have it in a spreadsheet.

```{r}
# your data
df_site <- tibble(ID = c('1','2','3'), 
                      Crop = c('Corn/Soy','Corn/Soy','Corn/Soy'),
                      Site = c('Arlington','Lancaster','Rock County'),
                      # Both coordinates in decimal format
                      latitude = c( 43.3091,42.8283,42.7255),
                      longitude = c(-89.3473,-90.7908,-89.0228),
                      Start = c('1988-01-01','1988-01-01','1988-01-01'),
                      End = c('2020-12-31','2020-12-31','2020-12-31'))

df_site <- df_site %>% 
  mutate_at(6:7, ymd) # as date

# this is for an exercise (intervals) - I am simulating a crop season to get GDD
df_time <- tibble(ID = c('1','2','3'), 
                      Crop = c('Corn/Soy','Corn/Soy','Corn/Soy'),
                      Site = c('Arlington','Lancaster','Rock County'),
                      # Both coordinates in decimal format
                      latitude = c( 43.3091,42.8283,42.7255),
                      longitude = c(-89.3473,-90.7908,-89.0228),
                      Start = c('2018-05-01','2018-06-01','2018-05-10'),
                      PRE = c('2018-05-02','2018-06-02','2018-05-12'),
                      POST = c('2018-06-03','2018-07-12','2018-06-15'),
                      End = c('2018-10-30','2018-11-15','2018-10-24'))

df_time <- df_time %>% 
  mutate_at(6:9, ymd) # as date

```

# Getting data

Read the pdf with description of all three sources of weather data. My understanding is that DayMet is the best.

## Daymet 

```{r constants}
# Constants for ET0 (Cobaner et al., 2017)
# Solar constant
Gsc = 0.0820 # (MJ m-2 min-1)
# Radiation adjustment coefficient (Samani, 2004)
kRs = 0.17
```


```{r function}
weather_daymet <- function(input, dpp = 0){ 
  input %>%
  mutate(Weather = pmap(list(ID = ID,
                             lat = latitude,
                             lon = longitude,
                             sta = Start - dpp,
                             end = End),
                        
        # Retrieving daymet data
        function(ID,lat,lon,sta,end){
                download_daymet(site = ID,
                                lat = lat, lon = lon,
                                # Extracting year
                                start = as.numeric(substr(sta,1,4)),
                                end = as.numeric(substr(end,1,4)),
                                internal = T, simplify = T)})) %>% 
  
  mutate(Weather = Weather %>% 
           
           map(~mutate(.,
                       Date = as_date(as.numeric(yday)-1, # Day of the year
                                         origin = paste0(year,'-01-01')),
                       Year = year(Date),
                       Month = month(Date),
                       Day = mday(Date)))  %>% 
           
           map(~dplyr::select(., yday, Year, Month, Day, Date,
                              measurement,value)) %>% 
           map(~spread(., 'measurement','value'))  %>% 
           map(~rename_all(., ~c("DOY", # Date as Day of the year
                                 "Year", # Year
                                 "Month", # Month 
                                 "Day", # Day of the month
                                 "Date", # Date as normal format
                                 "DL", # Day length (sec)
                                 "PP", # Precipitation (mm)
                                 "Rad", # Radiation (W/m2)
                                 "SWE", # Snow water (kg/m2)
                                 "Tmax", # Max. temp. (degC)
                                 "Tmin", # Min. temp. (degC)
                                 "VPD"))))  %>% # Vap Pres Def (Pa)
  
  mutate(Weather = pmap(list(sta=Start-dpp,
                             end = End,data=Weather), # Requested period
                        #~filter(..3, Date>=..1 & Date<= ..2))) %>% unnest() %>%
                        function(sta, end, data){
                          filter(data, Date >= sta & Date <= end) 
                        } )) %>% unnest(cols = c(Weather)) %>% 
  
  # Converting units or adding variables
  mutate(Rad = Rad*0.000001*DL, # Radiation (W/m2 to MJ/m2)
         Tmean = (Tmax+Tmin)/2, # Mean temperature (degC),
         VPD = VPD / 1000, # VPD (Pa to kPa),
         # Data for ET0
         lat_rad = latitude*0.0174533,
         dr = 1 + 0.033*cos((2*pi/365)*DOY),
         Sd = 0.409*sin((2*pi/365)*DOY - 1.39),
         ws = acos(-tan(lat_rad)*tan(Sd)),
         Ra = (24*60)/(pi) * Gsc * dr * (ws*sin(lat_rad)*sin(Sd)+
                                  cos(lat_rad)*sin(ws)),
         ET0_HS = 0.0135 * kRs * (Ra / 2.45) * (sqrt(Tmax-Tmin)) * (Tmean + 17.8),
         DL = (DL/60)/60 # Day length (hours)
          ) %>% dplyr::select(-lat_rad,-dr,-Sd,-ws,-Ra)
}
```



```{r get_data}
#dpp default = 0 - dpp = days prior start date

df_daymet <- weather_daymet(input = df_time, dpp = 0) 
```




```{r}
#df_daymet is ready - you have a full data set with weather data from the past 30-y
# There are many ways you can retrieve the data from it. Below just an simple plot
df_daymet %>% 
  mutate(Month = as_factor(Month)) %>% 
  ggplot(aes(x = DOY)) +
  geom_smooth(aes(y = Tmax), method = "loess", color = "red") +
  geom_smooth(aes(y = Tmin), method = "loess", color = "blue") +
  facet_grid(~ Site) 
```

## NASA-POWER 

```{r function-2}
weather_nasapower <- function(input, dpp = 0){
  input %>%
  # Retrieving the data from nasapower
  mutate(Weather = pmap(list(ID = ID,
                             lat = latitude,
                             lon = longitude,
                             sta = Start - dpp,
                             end = End),
          function(ID,lat,lon,sta,end){
          get_power(community = "AG",
                    dates = c(sta, end),
                    lonlat = c(lon, lat),
                    temporal_average = "DAILY",
                    # Variables (see package documents)
                    pars = c("T2M_MIN", # Min. temp. (degC)
                             "T2M_MAX", # Max temp. (degC)
                             "RH2M", # Relative Humidity 2M
                             "PRECTOT", # Precipitation (mm)
                             "ALLSKY_SFC_SW_DWN"))} ) )  %>% # Radiation (MJ/m2)

  mutate(Weather = Weather %>% 
           map(~as.data.frame(.)) %>%
           # Dates operations
           map(~mutate(., yday = lubridate::yday(YYYYMMDD),
                       Year = year(YYYYMMDD),
                       Month = month(YYYYMMDD),
                       Day = mday(YYYYMMDD)))  %>%
           map(~dplyr::select(., yday,Year, Month, Day,YYYYMMDD,
                              T2M_MIN,T2M_MAX,RH2M,
                              PRECTOT,ALLSKY_SFC_SW_DWN)) %>% 
           
           map(~rename_all(., ~c("DOY", # Day of the Year
                                 "Year", # Year
                                 "Month", # Month 
                                 "Day", # Day of the month
                                 "Date", # Date
                                 "Tmin", # Min. temp. (degC)
                                 "Tmax", # Max. temp. (degC)
                                 "RH", # Relative Humidity (%)
                                 "PP", # Precipitation (mm)
                                 "Rad")))) %>% # Radiation (MJ/m2)
  
  unnest(cols = c(Weather)) %>% ungroup() %>% 
  # Converting units or adding variables
  mutate(Tmean = (Tmax+Tmin)/2, # Mean temp. (degC)
         # Nasapower does not provide VPD values
         # However, it is possible to estimate it with Temp and RH.
         es = 0.6108 * exp((17.27*Tmean) / (Tmean+237.3)),
         ea = es * (RH / 100),
         # vapour Pressure deficit (kPa)
         VPD = es - ea,
         # Data for ET0
         lat_rad = latitude*0.0174533,
         dr = 1 + 0.033*cos((2*pi/365)*DOY),
         Sd = 0.409*sin((2*pi/365)*DOY - 1.39),
         ws = acos(-tan(lat_rad)*tan(Sd)),
         Ra = (24*60)/(pi) * Gsc * dr * (ws*sin(lat_rad)*sin(Sd)+
                                  cos(lat_rad)*sin(ws)),
         ET0_HS = 0.0135 * kRs * (Ra / 2.45) * (sqrt(Tmax-Tmin)) * (Tmean + 17.8)
         ) %>% dplyr::select(-es,-ea,-lat_rad,-dr,-Sd,-ws,-Ra)
}
```




```{r}
#dpp default = 0 - dpp = days prior start date

df_nasapower <- weather_nasapower(input = df_time, dpp = 0)
```

```{r}
df_nasapower %>% 
  mutate(Month = as_factor(Month)) %>% 
  ggplot(aes(x = DOY)) +
  geom_smooth(aes(y = Tmax), method = "loess", color = "red") +
  geom_smooth(aes(y = Tmin), method = "loess", color = "blue") +
  facet_grid(~ Site) 
```


## CHIRPS 

Here we download the daily-weather data from the CHIRPS database, and we process it to obtain common variables of agronomic value.

```{r function-3}
weather_chirps <- function(input, dpp = 0){ 
  input %>%
  # Retrieving the data from CHIRPS
  mutate(Weather = pmap(list(ID = ID,
                             lat = latitude,
                             lon = longitude,
                             sta = Start - dpp,
                             end = End),
                        function(ID, lat, lon, sta, end){
        get_chirps(data.frame(lon = c(lon), lat = c(lat)),
                   c(as.character(sta), as.character(end)))}) ,
  # Get prec. indices
        Indices = Weather %>% map(~precip_indices(., timeseries = TRUE, 
                                                   intervals = 30))) %>% 
  
  # Organizing dataframe
  mutate(Weather = Weather %>% 
           map(~as.data.frame(.)) %>% 
           map(~dplyr::select(., date, chirps)) %>%
           # Dates operations
           map(~mutate(., yday = lubridate::yday(date),
                       Year = year(date),
                       Month = month(date),
                       Day = mday(date))) %>% 
           map(~dplyr::select(., yday, Year, Month, Day, date,chirps)) %>%
           map(~rename_all(., ~c("DOY", "Year", "Month", "Day","Date", "PP"))),
         
         Indices = Indices %>% 
           map(~ as.data.frame(.)) %>% 
           map(~ spread(., 'index','value')) %>% 
           map(~ dplyr::select(., -id,-lon,-lat)) %>% 
           map(~rename(., Date = date))) %>% 
    mutate(Full = map2(.x = Weather, .y = Indices, ~ left_join(.x,.y))) %>% 
    dplyr::select(-Weather, -Indices) %>% 
    unnest(cols = c(Full))
}
```


```{r}
# Specify Days prior planting. Default is dpp = 0
# Lots of NA values - did not get this function well
df_chirps <- weather_chirps(input = df_time, dpp = 0)  
```


# Time intervals

## Season

```{r}
season <- df_time %>% 
  mutate(Intervals = # Create
           map2(.x=Start,.y=End,
                ~data.frame( # New data
                  Interval = c("Season"),
                  Start.in = c(.x),
                  End.in = c(.y) ) )) %>% 
  dplyr::select(ID, Site, Intervals) %>% 
  unnest(cols = c(Intervals))
```


## EVEN intervals

``` {r}
n = 4 # Number of intervals
dpp = 30 # Days prior planting

even <- df_time %>% 
  mutate(Intervals = # Create
      map2(.x=Start,.y=End,
      ~ tibble( # New data
      Interval = c("Prev", LETTERS[1:n+1]),
      Start.in = c(.x-dpp, seq.Date(.x,.y+1,length.out=n+1)[1:n]),
      End.in = c(.x-1, seq.Date(.x,.y+1,length.out=n+1)[2:(n+1)])))) %>% 
  dplyr::select(ID, Site, Intervals) %>% 
  unnest(cols = c(Intervals))
```


## CUSTOM intervals

Intervals like herbicide application as my example.

```{r}
# Counting # intervals
i = ncol(df_time[, 6:ncol(df_time)]) # Number of intervals


custom <- df_time %>% 
  mutate(Intervals = # Create
           pmap(list(x = Start - dpp,
                     y = Start,
                     z = PRE,
                     m = POST,
                     k = End),
                function(x,y,z,m,k){
                  data.frame( # New data
                  Interval = c(LETTERS[1:i]),
                  Name = c("Prev", "Plant-PRE","PRE-POST","POST-End"),
                  Start.in = c(x,y,z,m),
                  End.in = c(y-1,z-1,m-1,k) ) } ) ) %>% 
  
  dplyr::select(ID,Site,Intervals) %>% 
  unnest(cols = c(Intervals))
```

# SEASONAL SUMMARIES

## Summary function - DAYMET & NASA-POWER

```{r}
# Defining the function to summarize DAYMET and/or NASA-POWER
summary_daymet_nasapower <- function(input, intervals) {

  intervals %>% 
  
  # Merging weather data
  left_join(input %>% 
              # Nesting weather data back for each site-ID
  dplyr::select_if(
    names(.) %in% c("ID", "Crop", "Site", "Date","DL", "PP",
                    "Rad", "Tmax", "Tmin", "Tmean", "VPD", "ET0_HS")) %>% 
  group_by(ID,Crop,Site) %>% nest(.key = 'Weather') %>% ungroup()) %>% 
  
  mutate(Weather = pmap(list(x = Start.in,y = End.in, data = Weather),
                        function(x, y, data){
                          filter(data, Date >= x & Date < y)} ) )  %>% 
  
  mutate(Weather = Weather %>% # User must addapt depending on the crop
           map(~mutate(.,
                # Ext. Prec. event
                EPEi = case_when(PP>25~1, TRUE~0),
                # Ext.Temp. event
                ETEi = case_when(Tmax >= 30~1, TRUE~0), 
                # Tmax factor,  crop heat units (CHU)
                Ymax = case_when(Tmax < 10~0, 
                            TRUE ~ 3.33*(Tmax-10)-0.084*(Tmax-10)),
                # Tmin factor, Crop heat units (CHU)
                Ymin = case_when(Tmin<4.44~0, 
                            TRUE ~ 1.8*(Tmin-4.44)), 
                # Daily CHU
                Yavg = (Ymax+Ymin)/2,
                # Tmin threshold Growing Degrees.
                Gmin = case_when(Tmin >= 10 ~ Tmin, 
                            TRUE ~ 10),
                # Tmax threshold Growing Degrees.
                Gmax = case_when(Tmax <= 30 ~ Tmax, 
                            TRUE ~ 30),
                # Daily Growing Degree Units.
                GDU = ((Gmin + Gmax)/2) - 10
                       ) ) ) %>% 

# Summary for each variable

  mutate(# Duration of interval (days)
         Dur = Weather %>% map(~nrow(.)),
         # Accumulated PP (mm)
         PP = Weather %>% map(~sum(.$PP)),
         # Mean Temp (C)
         Tmean = Weather %>% map(~mean(.$Tmean)),
         # Accumulated Rad (MJ/m2)
         Rad = Weather %>% map(~sum(.$Rad)),
         # Accumulated VPD (kPa)
         VPD = Weather %>% map(~sum(.$VPD)),
         # Accumulated ET0 (mm)
         ET0_HS = Weather %>% map(~sum(.$ET0_HS)),
         # Number of ETE (#)
         ETE = Weather %>% map(~sum(.$ETEi)),
         # Number of EPE (#)
         EPE = Weather %>% map(~sum(.$EPEi)),
         # Accumulated Crop Heat Units (CHU)
         CHU = Weather %>% map(~sum(.$Yavg)),
         # Shannon Diversity Index for PP
         SDI = Weather %>% map(~diversity(.$PP, index="shannon")/
                                 log(length(.$PP))),
         # Accumulated Growing Degree Days (GDD) 
         GDD =  Weather %>% map(~sum(.$GDU))) %>% 
  
# Additional indices and final units
  dplyr::select(-Weather) %>% unnest() %>% 
  mutate(# Photo-thermal quotient (Q)
         Q_chu = Rad/CHU,
         Q_gdd = Rad/GDD,
         # Abundant and Well Distributed Water
         AWDR = PP*SDI) 
}
```

## Daymet summary
```{r}
# intervals = type of intervals (season, custom or even)
df_summary_daymet <- summary_daymet_nasapower(input = df_daymet,
                           intervals = season)
```

## NASA-POWER summary
```{r}
# intervals = type of intervals (season, custom or even)
df_summary_nasapower <-
  summary_daymet_nasapower(input = df_nasapower,
                           intervals = custom)
```

## Summary function - CHIRPS. <br/>
``` {r war}
#########

# Defining function to summarize CHIRPS data
summary_chirps <- function(input, intervals) {

intervals %>% 
  
  # Merging weather data
  left_join(input %>% 
              # Nesting weather data back for each site-ID
  dplyr::select(c(ID, Crop, Site, Date, PP)) %>% 
  group_by(ID,Crop,Site) %>% nest(.key = 'Weather') %>% ungroup()) %>% 
  
  mutate(Weather = pmap(list(x = Start.in,y = End.in, data = Weather),
                        function(x, y, data){
                          filter(data, Date >= x & Date < y)} ) )  %>% 
  
  mutate(Weather = Weather %>% # User must addapt depending on the crop
           map(~mutate(., EPEi = case_when(PP>25~1, TRUE~0) # Ext. Prec. event
                       ) ) )   %>%

# Summary for each variable
  
  mutate(# Duration of interval (days)
         Dur = Weather %>% map(~nrow(.)),
         # Accumulated PP (mm)
         PP = Weather %>% map(~sum(.$PP)),
         # Number of EPE (#)
         EPE = Weather %>% map(~sum(.$EPEi)) ,
         # Shannon Diversity Index for precipitation data
         SDI = Weather %>% map(~diversity(.$PP, index="shannon")/
                                 log(length(.$PP)))) %>% 
  
# Additional indices and final units
  dplyr::select(-Weather) %>% unnest() %>% 
  mutate(AWDR = PP*SDI) # Abundant and Well Distributed Water

}

```


## CHIRPS summary
```{r}
# intervals = type of intervals (season, custom or even)

df_summary_chirps <-
  summary_chirps(input = df_chirps,
                 intervals = custom)
```



# Historical

## Daymet 
```{r}
#dpp default = 0 - dpp = days prior start date

df_daymet_historical <- weather_daymet(input = df_site, dpp = 0) 
```
## NASA power

```{r}
#dpp default = 0 - dpp = days prior start date

df_nasapower_historical <- weather_nasapower(input = df_site, dpp = 0)
```


## Chirps

```{r}
# Specify Days prior planting. Default is dpp = 0
# Lots of NA values - did not get this function well
df_chirps_historical <- weather_chirps(input = df_site, dpp = 0)  
```


There are more codes in the "code_agromet_R.Rmd" to get historical intervals but I think you can get what you want using tidyverse functions (summarise, filter, mutate) from the histotical data set.



